---
title: "Sea Birds Analysis"
author: "Keegan S"
date: "13/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(here)
```

No assumptions had to be made thus far in data cleaning. I have simply extracted the data necessary to answer the brief questions.

First: import data.

```{r message=FALSE}
obs_clean <- read_csv(here::here("clean_data/obs_clean.csv"))
```


## Question 1: Which bird had the most individual sightings?

```{r}
#filter by count = 1, group by bird, count occurrences
ind_sight <- obs_clean %>%
  filter(count == 1) %>%
  group_by(common_name, scientific_name, abbreviation) %>%
  count() %>%
  rename("individual_sightings" = "n") %>%
  #arrange descending
  arrange(desc(individual_sightings)) 
#select and print topmost value
kable(ind_sight[1, ])
```

## Question 2: Which bird had the highest total count?

```{r}
#group by bird, summarise sum of count

highest_count <- obs_clean %>%
  group_by(common_name, scientific_name, abbreviation) %>%
  summarise("total_count" = sum(count)) %>%
  arrange(desc(total_count))

kable(highest_count[1, ])
```

## Question 3: Which bird had the highest total count above a latitude of -30?

```{r}
#filter by latitudes above -30, repeat total count from above

above_30 <- obs_clean %>%
  filter(lat > -30) %>%
  group_by(common_name, scientific_name, abbreviation) %>%
  summarise("total_count_above_30" = sum(count)) %>%
  arrange(desc(total_count_above_30))

kable(above_30[1, ])
```

## Question 4: How many different types of birds were only ever seen in groups of 1?

```{r}
#"group of 1" just means count of 1. No other alternative in the 
#data documentation. Filter by count = 1, count unique bird names.

groups_of_1 <- obs_clean %>%
  filter(count == 1) %>%
  #Note: Some of these are listed as "name" (unidentified). I am going to remove these, as it could artificially inflate the number of types. 
  #Remove unidentifieds with reg exp:
  filter(!str_detect(common_name, 
                     fixed("unidentified", ignore_case = TRUE)))
#print total number of unique entries
kable(length(unique(groups_of_1$common_name)))
```

## Question 5: How many penguins were seen? (Hint: there are many types of penguin)

```{r}
#will have to locate all instances of "penguin" in common_name, case-insensitive.

penguins <- obs_clean %>%
  filter(str_detect(common_name,
                    fixed("penguin", ignore_case = TRUE)))

#return sum of count of all penguin results:

kable(sum(penguins$count, na.rm = TRUE))
```

There were 158 penguins seen.
